### 2025-10-28 write-ups/manuscript/kid-tangrams-ms.Rmd
start ms

---
title: |
  | Preschoolers can form conventional pacts with each other 
  | to communicate about novel referents
bibliography: preschool-tangrams.bib
csl: apa7.csl
document-params: "10pt, letterpaper"

author-information: > 
    \author{Anonymized}

abstract: > 
  Learning language requires learning not only the content of language, but also how to use language to communicate. Iterated reference games provide a window into such skills, requiring rich communication as participants converge on mutually understandable names for initially novel referents. Some early experiments are interpreted as evidence that 4-5-year-old children cannot converge to the mutually understandable names needed to communicate in an iterated reference game. Here, we revisit young children's referential communicative abilities using a simpler, child-friendly paradigm. Across 51 pairs of children, we found that 4-5-year-olds successfully established reference with each other. Children were 85% accurate, and they often used descriptions similar to their partner's. These findings suggest that children’s capacity to construct effective referring expressions in novel contexts emerges earlier than once thought, consistent with the view that children show early pragmatic competence in supportive contexts.

keywords: >
    development; iterated reference game; pragmatics; preschoolers

output: cogsci2024::cogsci_paper
header-includes: \raggedbottom
#header-includes: \cogscifinalcopy
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(
  fig.width = 3, fig.height = 3, fig.crop = F,
  fig.pos = "tb", fig.path = "figs/",
  echo = F, warning = F, cache = F,
  message = F, sanitize = T
)

@@

expt_1 <- "data/expt_1"
expt_2 <- "data/expt_2"
mod_loc <- "code/models"
mod_results <- "code/models/summary"
mod_form <- "code/models/formulae"
time_palette <- c("practice" = "grey", "block 1" = "#E41A1C", "block 2" = "#377EB8", "block 3" = "#4DAF4A", "block 4" = "purple")

source(here("code/paper_helper.R"))
```

@@

```{r interface, fig.env = "figure*", fig.pos = "t!", out.width="\\textwidth", fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Experimental setup and procedure. Panel A shows the experimental setup. Panel B shows the 4 possible targets; names for targets are for cross-reference with later figures only. Panel C shows the procedure for Experiment 1; within critical blocks targets were ordered randomly. Panel D shows the procedure for Experiment 2. \\label{game}", cache=FALSE}

@@

More recently, though no evidence has directly contradicted @glucksberg1966, a number of other studies have revealed early emerging skills in the preschool years that support the use of referential communication. Preschool-aged children are faster to select a target when it is referred to in a consistent way [@graham2014; @matthews2010] and 6-year-olds are more likely to use consistent referential expressions when their partner is consistent [@koymen2014], suggesting that young children are sensitive to the norm of consistent descriptions in cooperative communication. Further, preschool-aged children adapt the informativeness of their referential expressions based on the visual content available to their interlocutor [@matthews2006;@nadig2002;@nilsen2009]. By age 5, children integrate information about other's perspectives into their comprehension of utterances [@sanjuan2015]. Overall, by late preschool, children can can reason about others’ perspectives to communicate more effectively.

@@

Given the task demands in @glucksberg1966 and work showing that children can succeed in a less demanding paradigm, here we revisit the question of child-child referential communication. In the present study, we re-examine young children's ability to establish effective referring expressions with each other in an iterated reference game using a simplified tablet-based paradigm. Across an initial study and a replication including a total of 51 pairs of 4-5-year-old children, we found that preschool-aged children were successful in an iterated reference game, suggesting that children's capacity to construct effective referring expressions in novel contexts emerges earlier than once claimed.

@@

4 and 5-year-old children were recruited from a university preschool during the school day. Children played with another child from the same class. Experiment 1 was conducted between June and August 2023. Pairs of children were included in analyses if they completed at least 8 of the 12 critical trials. We had 19 games that completed 12 critical trials, and 1 game that completed 11 critical trials. Of the 40 children, 21 were girls, and the median age was 57 months, with a range of 48-70 months.

@@

Children's selections and the time to selection were recorded from the experiment software. Children's descriptions were automatically transcribed from the video using Whisper [@radford2022] for the first pass and then hand-corrected by experimenters. Transcripts were hand-annotated for when each trial started, who said each line, and what referential descriptions were used. 
We excluded trials where the "teller" did not produce a description, or where all description was unintelligible and impossible to transcribe. After exclusions, we had `r clean_1 |> filter(!is.na(repNum)) |> nrow()` trials remaining.

Statistical analyses were run in brms [@burkner2018] with weakly informative priors. We report estimates and 95% credible intervals. The experimental set-up, analysis code, de-identified transcripts, and performance data for both experiments is available at [anonymized repo](https://osf.io/7d3ne/?view_only=d094ccd6438741bfbcaa7fde9fbb56b2).

@@

```{r fig-accuracy, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "Children's accuracy at selecting the correct target over time. Error bars are bootstrapped 95% CIs with a linear trend line overlaid. \\label{acc}" }
clean_2 |>
  filter(type != "practice") |>
  mutate(trial = trial - 3, expt = "Expt 2") |>
  bind_rows(clean_1 |> filter(type != "practice") |> mutate(trial = trial - 1, expt = "Expt 1")) |>
  ggplot(aes(x = trial, y = correct)) +
    geom_smooth(method = "lm", color = "black") +
  stat_summary(aes(color = type), fun.data = "mean_cl_boot") +
  #stat_summary(fun.data = "mean_cl_boot", geom = "linerange", color = "black") +
  scale_color_manual(values = time_palette) +
  # geom_bar(position = position_dodge(.9), color = "black", fill = "#19ADDE", width = .7) +
  # geom_text(aes(label = percent), hjust = 1.6, color = "white", size = 3) +
  xlab("Trial") +
  ylab("Accuracy") +
  geom_vline(xintercept = 4.5, lty="dashed")+
    geom_vline(xintercept = 8.5, lty="dashed")+
    geom_vline(xintercept = 12.5, lty="dashed")+
  scale_x_continuous(breaks = seq(1, 16, 1), expand = c(0.02, 0.02)) +
  scale_y_continuous(breaks = seq(0, 1, .2), expand = c(0, 0.04)) +
  coord_cartesian(ylim = c(0.3,1))+
  geom_hline(yintercept = .5) +
  facet_grid(expt ~ .) +
  theme(
    legend.position = "none",
    panel.grid = element_blank(),
    strip.background = element_blank()
  )
```

@@

Our primary measure of interest was whether children could accurately communicate the intended target. <!-- as prior work is often interpreted as indicating the children at kindergarten age cannot communicate about abstract shapes successfully [@glucksberg1966]. --> To test for changes in accuracy over time, we fit a Bayesian mixed effects logistic regression predicting accuracy.^[`r acc_form |> form()`] Children's accuracy was above chance (Odds Ratio: `r stats_text(acc_mod_1,1)`), and their accuracy slightly increased over the game (OR of one trial later: `r stats_text(acc_mod_1, 2)`, Figure \ref{acc}). This level of accuracy is generally in-line with accuracies from 4-year-olds playing with their parents in @leung2024 and indicates that children can understand and succeed at the task.

@@

```{r description, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "Length of description produced by the teller each trial. Grey dots are individual data points, colored dots are per trial means with bootstrapped 95% CIs. \\label{length}" }
clean_2 |>
  filter(type != "practice") |>
  mutate(trial = trial - 3, expt = "Expt 2") |>
  bind_rows(clean_1 |> filter(type != "practice") |> mutate(trial = trial - 1, expt = "Expt 1")) |>
  ggplot(aes(x = trial, y = words, color = type)) +
  geom_jitter(alpha = .5, color = "grey") +
  stat_summary(fun.data = "mean_cl_boot") +
  scale_x_continuous(breaks = seq(1, 20, 1), expand = c(0, 0)) +
    geom_vline(xintercept = 4.5, lty="dashed")+
    geom_vline(xintercept = 8.5, lty="dashed")+
    geom_vline(xintercept = 12.5, lty="dashed")+
  scale_color_manual(values = time_palette) +
  geom_smooth(method = "lm", color = "black") +
  xlab("Trial") +
  ylab("Words") +
  facet_grid(expt ~ .) +
  theme(
    legend.position = "none",
    panel.grid = element_blank(),
    strip.background = element_blank()
  )
```

@@

Some examples to illustrate the variety of effective descriptions children employed are shown in Table \ref{example} (these specific examples are from Experiment 2, but both experiments had similar distributions of descriptions).

@@

```{r sims, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "Semantic similarity between pairs of descriptions from different sources. Dots are means and lines are bootstrapped 95% CIs.  \\label{sim}" }
sims_2 |>
  mutate(source = case_when(
    speaker1 == speaker2 ~ "same teller",
    game1 == game2 ~ "same game",
    T ~ "different games"
  )) |>
  mutate(expt = "Expt 2") |>
  bind_rows(sims_1 |>
    mutate(source = case_when(
      speaker1 == speaker2 ~ "same teller",
      game1 == game2 ~ "same game",
      T ~ "different games"
    )) |> mutate(expt = "Expt 1")) |>
  ggplot(aes(x = source, y = sim, color = target1)) +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
  labs(y = "Cosine similarity", x = "Comparing two utterances from ...", color = "Target") +
  facet_grid(expt ~ .) +
  scale_color_solarized() +
  theme(
    legend.position = "bottom",
    panel.grid = element_blank(),
    strip.background = element_blank()
  )
```

@@

While description length is often used as a proxy for measuring convention formation, it does not capture semantic overlap between utterances. @boyce2024 introduced a more sensitive measure of semantic convergence that compares the content of utterances using word embeddings to trace how similarities within and across games change over time. <!--To measure convergence of utterances, they compared how the similarities between utterances from a given block and the final block (conventionalized) utterances increased over the course of the game. --> With only 3 blocks of descriptions, we do not expect semantic similarity for descriptions of a given target to show any meaningful change over time. However, we can test for a more coarse measure of sensitivity to partner: whether children's utterances are more like their partner's than children in other games'. If children are fully ego-centric [as suggested by @glucksberg1966], their choices of descriptions would be independent from their partners.

@@

We compared the semantic similarities between descriptions of the same target based on who produced the description (Figure \ref{sim}).  We used a Bayesian mixed effects linear regression to predict similarity.^[`r sim_across_between_form |> form()`] Utterances were more similar if they came from the same partnership (`r sim_across_between_1_results |> stats_text(2,3)`). Utterances were slightly more similar still if they came from the same person (`r sim_across_between_1_results |> stats_text(3,3)`), which is expected since children are likely to be fairly consistent with themselves. However, children used descriptions that were much more similar to their partner's than to other children's (Figure \ref{sim}), indicating sensitivity to their partner's expressions.

@@

@@

### Participants
Experiment 2 was run between March and August of 2024, at the same preschool as Experiment 1. No children participated in both experiments. 30 pairs of children completed all 16 critical trials, and 1 pair of children completed 10 critical trials. Our target age range was 4 and 5-year-olds, but one older 3-year-old was unintentionally included. Of the 62 children, 30 were girls, and the children had a median age of 56 months, and a range of 45-69 months.

@@

### Data processing
Data were processed in the same way as Experiment 1. After excluding trials where children did not give a description or where the experimenter echoed a child's description, we had `r clean_2 |> filter(!is.na(repNum)) |> nrow()` trials total.

@@

```{r sim2, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "Semantic similarity between descriptions from earlier blocks (1-3) and the last block in Experiment 2.  Heavy dots are means with bootstrapped 95% CIs; light dots are individual values. \\label{tolast}" }
sims_2 |>
  filter(game1 == game2) |>
  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
  filter(later == 4) |>
  mutate(same_speaker = ifelse(speaker1 == speaker2, "Same teller", "Different teller")) |>
  ggplot(aes(earlier, sim, color = target1)) +
    geom_smooth(color="black", method="lm")+
  geom_point(alpha = .1, position = position_dodge(width = .2)) +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
    #stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2), geom="line", lty="dashed") +
  #  geom_smooth( method="lm")+
  facet_wrap(~same_speaker) +
  scale_color_solarized() +
  scale_x_continuous(breaks=c(1,2,3), labels=c("1", "2", "3"))+
  labs(x = "Comparing an earlier block (1-3) to last block (4)", y = "Cosine Similarity", color = "Target") +
  theme(legend.position = "bottom",
            strip.background = element_blank(),
        #panel.grid=element_blank()
)

@@

<!--### Qualitative variation in interaction structure

-->

@@

@@

@@

@@

Prominent early studies claimed that young children cannot overcome their ego-centrism to coordinate with each other in reference games [@glucksberg1966]. However, more recent developmental work has found that young children show emerging communicative and pragmatic sensitivity, especially when cognitive demands are low. <!--Recent work has established that children are successful at producing and understanding referential expressions to novel shapes when playing reference games with their parents. -->Here, we revisited the question of preschoolers' performance in child-child reference games, using a scaffolded paradigm to reduce extraneous task demands.

>>>>

---
title             : |
  | Preschoolers can form conventional pacts with each other 
  | to communicate about novel referents

author: 
  - name          : "Veronica Boyce"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    email         : "vboyce@stanford.edu"
  - name          : "Robert Z. Sparks"
    affiliation   : "1"
  - name          : "Michael C. Frank"
    affiliation   : "1"

affiliation:
  - id            : "1"
    institution   : "Stanford University"

authornote: |
  TODO credit taxonomy etc

abstract: > 
  Learning language requires learning not only the content of language, but also how to use language to communicate. Iterated reference games provide a window into such skills, requiring rich communication as participants converge on mutually understandable names for initially novel referents. Some early experiments are interpreted as evidence that 4-5-year-old children cannot converge to the mutually understandable names needed to communicate in an iterated reference game. Here, we revisit young children's referential communicative abilities using a simpler, child-friendly paradigm. Across TODO pairs of children, we found that 4-5-year-olds successfully established reference with each other. Children were TODO% accurate, and they often used descriptions similar to their partner's. These findings suggest that children’s capacity to construct effective referring expressions in novel contexts emerges earlier than once thought, consistent with the view that children show early pragmatic competence in supportive contexts.
  <!-- https://tinyurl.com/ybremelq -->

keywords          : development; iterated reference game; pragmatics; preschoolers
wordcount         : "X"

bibliography      : preschool-tangrams.bib

floatsintext      : yes
linenumbers       : yes
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
csl               : apa7.csl
output            : papaja::apa6_pdf
header-includes   :
  - \usepackage{setspace}\singlespacing
---

```{r setup, include = FALSE}
library("papaja")

knitr::opts_chunk$set(
  fig.width = 3, fig.height = 3, fig.crop = F,
  fig.pos = "tb", fig.path = "figs/",
  echo = F, warning = F, cache = F,
  message = F, sanitize = T
)

@@

expt_1 <- "data/expt_1"
expt_2 <- "data/expt_2"
expt_3 <- "data/expt_3"
mod_loc <- "code/models"
mod_results <- "code/models/summary"
mod_form <- "code/models/formulae"
time_palette <- c("practice" = "grey", "block 1" = "#E41A1C", "block 2" = "#377EB8", "block 3" = "#4DAF4A", "block 4" = "purple")

library(ggtext)
labels <- tribble(~target, ~img,
                  "hold", "I1.jpg",
                  "swim", "D1.jpg",
                  "walk", "B1.jpg",
                  "jump", "E1.jpg",
                  "baby","tangram_H.png",
                  "bunny","tangram_E.png",
                  "dancer","tangram_G.png",
                  "zombie", "tangram_A.png"
                  ) |> 
  mutate(img_source= str_c("<img src='",img,"' width='20' />"))

#source(here("code/paper_helper.R"))
joint_data <- read_csv(here("data/processed_for_paper/joint_with_demog.csv"))
joint_similarities <-   read_csv(here("data/processed_for_paper/sims.csv"))

### functions for printing things nicely!
stats <- function(model, row, decimal = 2) {
  model <- model |>
    mutate(
      Estimate = round(Estimate, digits = decimal),
      Lower = round(lower, digits = decimal),
      Upper = round(upper, digits = decimal),
      `Credible Interval` = str_c("[", Lower, ", ", Upper, "]")
    ) |>
    select(Term, Estimate, `Credible Interval`)
  str_c(model[row, 1], ": ", model[row, 2], " ", model[row, 3])
}

stats_text <- function(model, row, decimal = 2) {
  model <- model |>
    mutate(
      Estimate = round(Estimate, digits = decimal) |> formatC(format='f', digits=decimal ),
      Lower = round(lower, digits = decimal) |> formatC(format='f', digits=decimal ),
      Upper = round(upper, digits = decimal) |> formatC(format='f', digits=decimal ),
      `Credible Interval` = str_c("[", Lower, ", ", Upper, "]")
    ) |>
    select(Term, Estimate, `Credible Interval`)
  str_c(model[row, 2], "  ", model[row, 3])
}

form <- function(model_form) {
  dep <- as.character(model_form$formula[2])
  ind <- as.character(model_form$formula[3])

  str_c(dep, " ~ ", ind) |>
    str_replace_all(" ", "") |>
    str_replace_all("\\*", " $\\\\times$ ") |>
    str_replace_all("\\+", "&nbsp;+ ") |>
    str_replace_all("~", "$\\\\sim$ ")
}
```

TODO all figures could be smaller!

@@

```{r interface, fig.env = "figure", fig.pos = "t!", out.width="\\textwidth", fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Experimental setup and procedure. Panel A shows the experimental setup. Panel B shows the 4 possible targets; names for targets are for cross-reference with later figures only. Panel C shows the procedure for Experiment 1; within critical blocks targets were ordered randomly. Panel D shows the procedure for Experiment 2. \\label{game}", cache=FALSE}

@@

TODO review the literature for 
* any other studies we should include
* any studies we want to say more about than we had room for in cogsci

More recently, though no evidence has directly contradicted @glucksberg1966, a number of other studies have revealed early emerging skills in the preschool years that support the use of referential communication. Preschool-aged children are faster to select a target when it is referred to in a consistent way [@graham2014; @matthews2010] and 6-year-olds are more likely to use consistent referential expressions when their partner is consistent [@koymen2014], suggesting that young children are sensitive to the norm of consistent descriptions in cooperative communication. Further, preschool-aged children adapt the informativeness of their referential expressions based on the visual content available to their interlocutor [@matthews2006;@nadig2002;@nilsen2009]. By age 5, children integrate information about other's perspectives into their comprehension of utterances [@sanjuan2015]. Overall, by late preschool, children can can reason about others’ perspectives to communicate more effectively. 
* [@sanjuan2025] When 4-5 year old children learn that some objects are only visible to them and not their partner, they are faster to rule out these competitors and select a target than when they believe the interlocuter is aware of the competitors as well. 
* [@nilsen2009] take into account partners perspective for both speaker and listening (4-5 yos); 3-4 year olds also on listening only; relate this to inhibitory control.

@@

- Bohn, Kachel, Tomasello (2019) Young Children spontaneously recreate core properties of language in a new modality

    - 4, 6, (and 8 in a separate thing) communicating target image via video link (no audio after practice trials)

    - comprehension good and increasing, gestures used back by second partner when roles switch

    - (out of 5 pictures, including “blank”)

    - do see more similaritiy w/i dyads especially for 6 versus 4 yos

May want to set up for change over age? And a little more on why 4-5 is a good age -- it's a time of rapid development, we think from prior work that children can comprehend the task. Just before big changes in linguistic input from school + reading. ...

Given the task demands in @glucksberg1966 and work showing that children can succeed in a less demanding paradigm, here we revisit the question of child-child referential communication. In the present study, we re-examine young children's ability to establish effective referring expressions with each other in an iterated reference game using a simplified tablet-based paradigm. Across three studies including a total of ZZ pairs of 4-5-year-old children, we found that preschool-aged children were successful in an iterated reference game, suggesting that children's capacity to construct effective referring expressions in novel contexts emerges earlier than once claimed.

@@

4 and 5-year-old children were recruited from a university preschool during the school day. Children played with another child from the same class. Experiment 1 was conducted between June and August 2023. Pairs of children were included in analyses if they completed at least 8 of the 12 critical trials. We had 19 games that completed 12 critical trials, and 1 game that completed 11 critical trials. Of the 40 children, 21 were girls, and the median age was 57 months, with a range of 48-70 months. An additional XX games were excluded for being pilot runs or not completing enough trials.

@@

Children's selections and the time to selection were recorded from the experiment software. Children's descriptions were automatically transcribed from the video using Whisper [@radford2022] for the first pass and then hand-corrected by experimenters. Transcripts were hand-annotated for when each trial started, who said each line, and what referential descriptions were used. 
We excluded trials where the "teller" did not produce a description, or where all description was unintelligible and impossible to transcribe. After exclusions, we had `r joint_data |> filter(expt=="1") |> nrow()` trials remaining.

Statistical analyses were run in brms [@burkner2018] with weakly informative priors. We report estimates and 95% credible intervals. The experimental set-up, analysis code, de-identified transcripts, and performance data for all experiments is available at [anonymized repo](https://osf.io/7d3ne/?view_only=d094ccd6438741bfbcaa7fde9fbb56b2).

@@

```{r fig-accuracy, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "Children's accuracy at selecting the correct target over time. Error bars are bootstrapped 95% CIs with a linear trend line overlaid. \\label{acc}" }
joint_data |>
  ggplot(aes(x = trial+1, y = correct)) +
    geom_smooth(method = "lm", color = "black") +
  stat_summary(aes(color = type), fun.data = "mean_cl_boot") +
  #stat_summary(fun.data = "mean_cl_boot", geom = "linerange", color = "black") +
  scale_color_manual(values = time_palette) +
  # geom_bar(position = position_dodge(.9), color = "black", fill = "#19ADDE", width = .7) +
  # geom_text(aes(label = percent), hjust = 1.6, color = "white", size = 3) +
  labs(color="Block")+
  xlab("Trial") +
  ylab("Accuracy") +
  geom_vline(xintercept = 4.5, lty="dashed")+
    geom_vline(xintercept = 8.5, lty="dashed")+
    geom_vline(xintercept = 12.5, lty="dashed")+
  scale_x_continuous(breaks = seq(1, 16, 1), expand = c(0.02, 0.02)) +
  scale_y_continuous(breaks = seq(0, 1, .2), expand = c(0, 0.04)) +
  coord_cartesian(ylim = c(0.3,1))+
  geom_hline(yintercept = .5) +
  facet_grid(str_c("Expt ",expt) ~ .) +
  theme(
    panel.grid = element_blank(),
    strip.background = element_blank()
  )
```

@@

Our primary measure of interest was whether children could accurately communicate the intended target, as prior work is often interpreted as indicating the children at kindergarten age cannot communicate about abstract shapes successfully [@glucksberg1966]. To test for changes in accuracy over time, we fit a Bayesian mixed effects logistic regression predicting accuracy.^[`r acc_form |> form()`] Children's accuracy was above chance (Odds Ratio: `r stats_text(acc_mod_1,1)`), and their accuracy slightly increased over the game (OR of one trial later: `r stats_text(acc_mod_1, 2)`, Figure \ref{acc}). This level of accuracy is generally in-line with accuracies from 4-year-olds playing with their parents in @leung2024 and indicates that children can understand and succeed at the task.

@@

```{r description, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "Length of description produced by the teller each trial. Grey dots are individual data points, colored dots are per trial means with bootstrapped 95% CIs. \\label{length}" }
joint_data |> 
  ggplot(aes(x = trial+1, y = words, color = type)) +
  geom_jitter(alpha = .5, color = "grey") +
  stat_summary(fun.data = "mean_cl_boot") +
  scale_x_continuous(breaks = seq(1, 20, 1), expand = c(0, 0)) +
    geom_vline(xintercept = 4.5, lty="dashed")+
    geom_vline(xintercept = 8.5, lty="dashed")+
    geom_vline(xintercept = 12.5, lty="dashed")+
  scale_color_manual(values = time_palette) +
  geom_smooth(method = "lm", color = "black") +
  xlab("Trial") +
  ylab("Words") +
  facet_grid(str_c("Expt ",expt) ~ .) +
  theme(
    legend.position = "none",
    panel.grid = element_blank(),
    strip.background = element_blank()
  )
```

@@

Some examples to illustrate the variety of effective descriptions children employed are shown in Table \ref{example} (these specific examples are from Experiment 2, but both experiments had similar distributions of descriptions). TODO add examples from expt 3 as well??

@@

```{r sims, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "Semantic similarity between pairs of descriptions from different sources. Dots are means and lines are bootstrapped 95% CIs.  \\label{sim}" }

joint_similarities|>
  mutate(source = case_when(
    speaker1 == speaker2 ~ "same teller",
    game1 == game2 ~ "same game",
    T ~ "different games")) |> 
  left_join(labels) |> 
  ggplot(aes(x = source, y = sim, color = img_source)) +
      stat_summary(aes(group=img_source),fun.data = "mean_cl_boot", position = position_dodge(width = .3), geom="line") +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .3)) +
  labs(y = "Cosine similarity", x = "Comparing two utterances from ...", color = "Target") +
  facet_grid(str_c("Expt ",expt) ~ .) +
  scale_color_brewer(palette="Dark2") +
  theme(
    legend.position = "right",
    legend.text = element_markdown(color = "black", size = 11),
    panel.grid = element_blank(),
    strip.background = element_blank()
  )
```

@@

While description length is often used as a proxy for measuring convention formation, it does not capture semantic overlap between utterances. @boyce2024 introduced a more sensitive measure of semantic convergence that compares the content of utterances using word embeddings to trace how similarities within and across games change over time. To measure convergence of utterances, they compared how the similarities between utterances from a given block and the final block (conventionalized) utterances increased over the course of the game. With only 3 blocks of descriptions, we do not expect semantic similarity for descriptions of a given target to show any meaningful change over time. However, we can test for a more coarse measure of sensitivity to partner: whether children's utterances are more like their partner's than children in other games'. If children are fully ego-centric [as suggested by @glucksberg1966], their choices of descriptions would be independent from their partners.

@@

We compared the semantic similarities between descriptions of the same target based on who produced the description (Figure \ref{sim}).  We used a Bayesian mixed effects linear regression to predict similarity.^[`r sim_across_between_form |> form()`] Utterances were more similar if they came from the same partnership (`r sim_across_between_1_results |> stats_text(2,3)`). Utterances were slightly more similar still if they came from the same person (`r sim_across_between_1_results |> stats_text(3,3)`), which is expected since children are likely to be fairly consistent with themselves. However, children used descriptions that were much more similar to their partner's than to other children's (Figure \ref{sim}), indicating sensitivity to their partner's expressions. 
%TODO consider dropping this because it's not that strong an analysis?

@@

TODO UPDATE EXAMPLES TO HAVE SOME FROM EXPT 3

@@

### Participants
Experiment 2 was run between March and August of 2024, at the same preschool as Experiment 1. No children participated in both experiments. 30 pairs of children completed all 16 critical trials, and 1 pair of children completed 10 critical trials. Our target age range was 4 and 5-year-olds, but one older 3-year-old was unintentionally included. Of the 62 children, 30 were girls, and the children had a median age of 56 months, and a range of 45-69 months. An additional XX games were excluded for being pilot runs or not completing enough trials.

@@

### Data processing
Data were processed in the same way as Experiment 1. After excluding trials where children did not give a description or where the experimenter echoed a child's description, we had `r joint_data |> filter(expt=="2") |> nrow()` trials total.

@@

```{r sim2, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "Semantic similarity between descriptions from earlier blocks (1-3) and the last block in Experiments 2 and 3.  Heavy dots are means with bootstrapped 95% CIs; light dots are individual values. \\label{tolast}" }
joint_similarities |> 
  filter(game1==game2) |> 
  filter(later==3) |> 
  mutate(same_speaker = ifelse(speaker1 == speaker2, "Same teller", "Different teller")) |>left_join(labels) |> 
  ggplot(aes(earlier+1, sim, color = img_source)) +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
    stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2), geom="line") +
    geom_smooth(color="black", method="lm")+
  facet_grid(str_c("Expt ", expt)~same_speaker) +
  scale_x_continuous(breaks=c(1,2,3), labels=c("1", "2", "3"))+
  labs(x = "Comparing an earlier block (1-3) to last block (4)", y = "Cosine Similarity", color = "Target") +
scale_color_brewer(palette="Dark2") +
  theme(
    legend.position = "right",
    legend.text = element_markdown(color = "black", size = 11),
    panel.grid = element_blank(),
    strip.background = element_blank()
  )

@@

# Experiment 3

The first two experiments demonstrated children's ability to communicate referentially, but were limited in their generality by the use of a specific set of 4 fairly humanoid tangram figures. To explore the generality of the results to a different set of potentially harder tangram stimuli, we conducted Experiment 3.

## Methods
As Experiment 2 was very similar to Experiment 2, we focus on the changes made compared to Experiment 2. Experiment 3 was pre-registered at TODO [anonymized link].

The procedure for Experiment 3 matched Experiment 2, and the only change was in the stimuli used.

### Participants
Experiment 3 was run between XX and YY , at the same preschool as Experiments 1 and 2. Because the stimuli changed, we allowed children who had participated in a prior experiment to participate again, and ZZ participants in experiment 2 were also in experiment 3. TODO pairs of children completed all 16 critical trials, and TODO. Our target age range was 4 and 5-year-olds. Of the TODO children, QQ were girls, and the children had a median age of ZZ months, and a range of XX-YY months. An additional XX games were excluded for being pilot runs, including repeat participants, not completing enough trials, or including a 3-year-old participant.

### Materials
A different set of critical images were used (see Figure TODO). These were chosen from a set of commonly used tangram images (CITATIONS), to be fairly distinct from each other. We also switched the stimuli for the 3rd and 4th practice trials to be more abstract, so children would not be surprised by the critical stimuli not having convention names (TODO figure).

### Procedure
The procedure was the same as for Experiment 2. The bubble popping practice was modified so that both rounds of bubbles were colored dots, because children in Experiment 2 sometimes did not identify the smiley faces as more bubbles to pop.

### Data processing
Data were processed in the same way as Experiment 1. After excluding trials where children did not give a description or where the experimenter echoed a child's description, we had TODO trials total.

## Results

### Accuracy and speed

@@

## Pooled analysis

@@

## Sensitivity to prior result

@@

## Age and gender as moderators of referential success

We did not account for children's age and gender when pairing children to play the game, so children of different ages (within the 4-5 year old target age range) sometimes played together, and children sometimes played with same-gender and sometimes with different-gender partners. As children of this age are rapidly developing their linguistic skills (CITATION NEEDED), we looked at possible effects of children's age and gender on performance as an exploratory measure.

```{r age-acc, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "Accuracy of selections across all experiments, graphed by the mean age of child in that game, and colored by the number of girls in the pair. Each dot is a game, and linear fits are overlayed. \\label{agewords}" }
joint_data |>
  mutate(mean_age = (listener_age + speaker_age) / 2, girlness = (listener_gender == "female") + (speaker_gender == "female")) |>
  ggplot(aes(x = mean_age, y = correct, color = as.character(girlness))) +
  labs(x="Mean age of participants (months)", y="Accuracy", color="Number of girls")+
  geom_jitter(data=joint_data|>
  mutate(mean_age = (listener_age + speaker_age) / 2, girlness = (listener_gender == "female") + (speaker_gender == "female")) |> group_by(expt, game, mean_age, girlness) |> summarize(correct=mean(correct, na.rm=T))) +
  scale_color_manual(values=c("#DC322FFF","#2AA198FF" ,"#268BD2FF"  ))+
  geom_smooth(method = "lm", show.legend=F)
```

```{r age-words, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "Number of words of description produced by the teller across all experiments, graphed by the age of teller in that game, and colored by teller gender. Each dot is a game, and linear fits are overlayed. \\label{ageacc}" }
joint_data |>  ggplot(aes(x = speaker_age, y = words, color = speaker_gender)) +
  geom_smooth(method = "lm", show.legend=F)+
  geom_jitter(data=joint_data |> group_by(speaker_age, speaker, expt, game, speaker_gender) |> summarize(words=mean(words, na.rm=T)))+
  scale_color_manual(values=c("#268BD2FF", "#DC322FFF"))+
  labs(x="Teller Age (months)", y="Words", color="Teller gender")
```

For accuracy ... TODO

For description length ... TODO

These findings of a slight effect of age and a slight female-advantage are consistent with a wider literature of ...

# Qualitative variation in interaction structure

* maybe cut this section
* do want to make the point that there's really wide variability (consistent with older kids and also adults for that matter) -- some pairs one kid does the describing including of options and the other chooses

* may want to be clear that structuring the interaction in an effective way is a skill that not all the kiddos have -- some relied more on the experimenter saying whose turn and needed more prompting with "do you want to give more hints" or "do you want to ask a question". whereas other groups rapidly produced descriptions and then either clicked or asked for (and got) clarification (either completely on their own, or after the option was questions offered as a possibility). 
that said, even when children were struggling, some of their attempts such as gesturing or turning ipads or asking (can you show me) are indicative of attempts at conversational monitoring and solving the communication problem (albeit not in ways we that followed the "rules" of the game)
where we see what looks like failures it may be hard to tell what part isn't working (is it knowing what i know, or what other knows, or cognitive control, or producing questions, or sampling)

# (unclear if include)
could also do some qualitative comparison of the *types* of descriptions that kiddos give versus kid-parent vs adult ?

@@

Prominent early studies claimed that young children cannot overcome their ego-centrism to coordinate with each other in reference games [@glucksberg1966]. However, more recent developmental work has found that young children show emerging communicative and pragmatic sensitivity, especially when cognitive demands are low. Recent work has established that children are successful at producing and understanding referential expressions to novel shapes when playing reference games with their parents. Here, we revisited the question of preschoolers' performance in child-child reference games, using a scaffolded paradigm to reduce extraneous task demands.

Across 3 experiments and TODO pairs of 4 and 5-year-old children, we tested how well children could produce referential expressions that allowed their partner to find a matching abstract shape. Children varied substantially in what sorts of descriptions they produced, but overall accuracy was high (TODO%), indicating that children were generally able to produce adequate descriptions. Additionally, children's utterances showed signs of converging toward conceptual pacts. While this task is substantially scaled down relative to measures used for adult competence, it does suggest that the relevant communication skills are present at least in rudimentary form by the end of the preschool years.
